---
title: "p8105_hw3_gl2458"
author: "Rachel Lee"
date: "10/14/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tibble)
```

# PROBLEM 1

```{r}
library(p8105.datasets)
data("instacart")

n_order_per_hour_of_day = 
  instacart %>% 
  group_by(order_hour_of_day) %>%
  distinct(order_id) %>% 
  summarise(n_order_per_hour = n()) %>% 
  arrange(desc(n_order_per_hour))
```

## Description

The instacart dataset contains `r ncol(instacart)` variables and `r nrow(instacart)` observations. The key variables are `aisle` and `aisle_id`, and they represent the names and identifiers of distinct aisles, respectively. `product_name` and `product_id` denote names and identifiers of distinct products (items), respectively.`order_dow` and `order_hour_of_day` represent the day of the week and the hour of the day that the order was placed, respectively. Distinct products purchased within the same order are listed in separate rows. 

## Problem 1.1

How many aisles are there, and which aisles are the most items ordered from?

```{r prob1.1_instacart}
instacart %>% 
  distinct(aisle) %>% 
  nrow()
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

There are `r instacart %>% distinct(aisle) %>% nrow()` aisles. Each observation represents individual items purchased. Hence, the number of observations by aisle is equivalent to the number of items ordered from each aisle. By arranging the observations by aisle, we can observe that the most items are ordered from the `fresh vegetables` aisle, followed by `fresh fruits` aisle and `packaged vegetables fruits` aisle.

## Problem 1.2

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r prob1.2_instacart}
instacart %>% 
  filter(aisle == "asian foods") %>% 
  select(order_id, aisle)
instacart %>% 
  group_by(aisle) %>% 
  summarise(n_items = n()) %>% 
  filter(n_items > 10000) %>% 
  arrange(desc(n_items))%>%
  ggplot(aes(x = aisle, y = n_items)) +
  geom_point() +
  scale_y_continuous(breaks = c(10000, 50000, 100000, 150000),
                    limits = c(10000, 200000), 
                    labels = c("10000", "50000", "100000", "150000")
                     ) +
  theme(axis.text.x = element_text(color = "#666666", angle = 90, size = rel(0.8), hjust = 1)) +
  labs(
    title = "Number of items purchased from each aisle",
    x = "Aisles",
    y = "Number of items", 
    caption = "Instacart dataset"
  ) 
```

The two aisles with most orders are 1) fresh vegetables and 2) fresh fruits. 

## Problem 1.3

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r prob1.3_instacart}
instacart %>% 
  group_by(aisle, product_name) %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>%  
  summarise(pop_product = n()) %>% 
  group_by(aisle) %>% 
  filter(pop_product == max(pop_product)) %>% 
  knitr::kable(col.names = c("Aisle", "Product", "Amount Sold"))
```

By grouping aisles and products, we can count the number of each product sold per aisle. Then, we can group it by aisle to observe the most popular product sold in each aisle. 

Most popular products in the following aisles are: 
  * baking ingredients - Light Brown Sugar
  * dog food care - Snacks Sticks Chicken and Reci Recipe Dog Treats
  * packaged vegetables fruits - Organic Baby Spinach
  
## Problem 1.4

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r prob1.4_instacart}
instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  mutate(order_dow = recode(order_dow, '0' = "Sunday", '1' = "Monday", '2' = "Tuesday", '3' = "Wednesday", '4' = "Thursday", '5' = "Friday", '6' = "Saturday")) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(mean_hour_of_day = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour_of_day) %>% 
  knitr::kable(digits = 2)
```

# Problem 2

Importing and cleaning the data


```{r prob2_dataimport}
brfss = 
  p8105.datasets::brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(resp_id = respid, location_abbr = locationabbr, location_desc = locationdesc) %>% #further manual renaming to appropriate variable names
  filter(topic == "Overall Health") %>% 
  select(year, location_abbr, location_desc, response, data_value)
brfss$response = 
  brfss$response %>%  
  factor(levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))
brfss
```

## Problem 2.1 

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r prob2.1_brfss}
brfss %>% 
  ungroup() %>% 
  filter(year %in% c(2002)) %>% 
  distinct(location_abbr, location_desc, year) %>% 
  group_by(location_abbr, year) %>% 
  count() %>% 
  filter(n >= 7) %>% 
  select(location_abbr, year) %>% 
  knitr::kable()
# for year 2010
brfss %>% 
  ungroup() %>% 
  filter(year %in% c(2010)) %>% 
  distinct(location_abbr, location_desc, year) %>% 
  group_by(location_abbr, year) %>% 
  count() %>% 
  filter(n >= 7) %>% 
  select(location_abbr, year) %>% 
  knitr::kable()
```

## Problem 2.2 

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).


```{r prob2.2_brfss}
brfss %>% 
  filter(response == "Excellent") %>% 
  group_by(location_abbr, year) %>% 
  summarise(mean_data_value = mean(data_value, na.rm = TRUE) 
  ) %>% 
  ggplot(aes(x = year, y = mean_data_value, color = location_abbr, group = location_abbr)) + geom_line() +
  labs(
      title = "Distribution of state-level averages for each year",
      x = "Year",
      y = "Mean Data value"
    )  +
  theme_bw()
``` 

## Problem 2.3

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r prob2.3_brfss}
brfss %>%
  ungroup() %>% 
  filter(year %in% c(2006, 2010) & location_abbr == "NY") %>%
  ggplot(aes(x = response, y = data_value, color = response)) + geom_point() + facet_grid(~ year) +
   labs(
      title = "Distribution of data value for each response in NY state",
      x = "Response",
      y = "Data value"
    ) 
```

# Problem 3

## Problem 3.1 

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r prob3.1_accel}
accel = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute", 
    names_prefix = "activity_",
    values_to = "activity") %>% 
  mutate(day_type = if_else(
  day %in% c("Saturday", "Sunday"),"weekend", "weekday"),
  day_type = as.factor(day_type),
  minute = as.numeric(minute),
  activity = as.numeric(activity), 
  minute = as.numeric(minute),
  day = as.factor(day), 
  week = as.factor(week),
  day_id = as.numeric(day_id))
```

### Description

There are **`r nrow(accel)` observations** and **`r ncol(accel)` variables**.

The variables in the dataset are **`r colnames(accel)`**.

## Problem 3.2

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r prob3.2_accel}
accel %>% 
  group_by(week, day) %>% 
  summarise(
    total_activity = sum(activity)
  ) %>% 
  knitr::kable(col.names = c("Week", "Day", "Total Activity"))
```

## Problem 3.3

Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.


```{r prob3.3_accel}
accel %>% 
  mutate(hour = minute %/% 60) %>%
  group_by(day, hour) %>%
  summarise(
    total_activity = sum(activity)
  ) %>%
  ggplot(aes(x = hour, y = total_activity, color = day)) + 
    geom_line() + 
    theme(legend.position = "bottom") + 
  scale_x_continuous(breaks = c(0:24)) +
  scale_y_continuous(
    breaks = c(50000, 150000, 250000)
  ) + 
  labs(
    x = "Hour",
    y = "Total Activity", 
    title = "24-hour activity time courses for each day"
  )
```

The peak activity occurs on fridays between 19th and 21st hour. The activity is at the lowest between 0th to 4th hour in the morning. The trend shows the activity increases after.



